Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-07-27 Paper/assa.tex
..

\subsection{Active subspaces}
\label{sub:ac}
%We recall some background regarding active subspaces~\cite{Constantine:2015}.
As mentioned earlier, the active subspace is low-dimensional linear subspace that constitutes
important directions in a model's input space~\cite{Constantine:2015}. The effective variability
in the model output due to uncertain inputs is predominantly captured along these directions.
Mathematically, these directions are essentially the eigenvectors of a positive semi-definite
matrix, $\mat{C}$, defined as follows:
%
\be
\mat{C} = \int_\Omega (\nabla_{\bm{\xi}}f)(\nabla_{\bm{\xi}}f)^\top dP_{\bm{\xi}}
\label{eq:C}
\ee
%
where $\Omega$:~$[-1,1]^{N_p}$ is the domain of $\bm{\xi}$, and $N_p$ is the number of uncertain
inputs. $dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous and differentiable
in the considered parameter domain. Hence, it is possible that a given $f(\bm{\xi})$ might not admit an
active subspace. However, it is of remarkable interest to investigate if one exists since the subspace could
be exploited to reduce the dimensionality of the problem and hence the associated computational effort.
$\nabla_{\bm{\xi}}f$ denotes a vector with individual components
being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. Since $\mat{C}$ is symmetric and
positive semi-definite, it admits a real eigenvalue decomposition:
%
\be
\mat{C} = \bm{W}\bm{\Lambda}\bm{W}^\top
\ee
%
where $\bm{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with $\lambda_i$'s
 sorted in descending order, $i:\{1,2,\ldots,N_p\}$. The eigenpairs are partitioned about
 the $i^{\text{th}}$ eigenvalue such that the ratio, $\left(\frac{\lambda_i}{\lambda_{i+1}}\right)$ $\gg$ 1
 as follows:
 %
 \be
 \bm{W} = [\bm{W}_1~\bm{W}_2],~~\bm{\Lambda} = \begin{bmatrix}\bm{\Lambda}_1 & \\  &
  \bm{\Lambda}_2 \\
\end{bmatrix}
 \ee
 %
 where $\bm{W}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as the active subspace,
 and $\bm{\Lambda}_1$ is the corresponding set of eigenvalues. The active subspace is used to 
 construct the so-called active variables, $\bm{w}_1^\top \bm{\xi}$ and $f(\bm{\xi}(\theta))$ is transformed
 as $G(\bm{w}_1^\top \bm{\xi})$. In practice, the set of samples, $\bm{\xi}_i$'s are projected to the
 physical space as $\bm{\theta}_i$'s at which the model output, $f(\bm{\xi}(\theta))$ is computed. 
 The matrix, $\mat{C}$ is typically approximated using pseudo-random sampling techniques such 
 as Monte Carlo and Latin hypercube sampling. The integral in~\eqref{eq:C} is replaced with a
 summation as follows:
 %
 \be
 \mat{C}\approx \hat{\mat{C}} = \frac{1}{N}\sum\limits_{i=1}^{N} 
 (\nabla_{\bm{\xi}}f(\bm{\xi}_i))(\nabla_{\bm{\xi}}f(\bm{\xi}_i))^\top
 = \hat{\bm{W}}\hat{\bm{\Lambda}}\hat{\bm{W}}^\top
 \ee
 %
 It has been shown that the accuracy of approximated dominant eigenspace, $\hat{\bm{W}}_1$ is inversely
 proportional to the difference between the smallest eigenvalue in $\hat{\bm{\Lambda}}_1$ and the
 largest eigenvalue in $\hat{\bm{\Lambda}}_2$~\cite{Constantine:2014}.

>>>>

\subsection{Active subspaces}
\label{sub:ac}
%We recall some background regarding active subspaces~\cite{Constantine:2015}.
As mentioned earlier, the active subspace is low-dimensional linear subspace that constitutes
important directions in a model's input space~\cite{Constantine:2015}. The effective variability
in the model output due to uncertain inputs is predominantly captured along these directions.
Mathematically, these directions are essentially the eigenvectors of a positive semi-definite
matrix, $\mat{C}$, defined as follows:
%
\be
\mat{C} = \int_\Omega (\nabla_{\bm{\xi}}f)(\nabla_{\bm{\xi}}f)^\top dP_{\bm{\xi}}
\label{eq:C}
\ee
%
where $\Omega$:~$[-1,1]^{N_p}$ is the domain of $\bm{\xi}$, and $N_p$ is the number of uncertain
inputs. $dP_{\bm{\xi}}$ = $p(\bm{\xi})d\bm{\xi}$, where $p(\bm{\xi})$ is the joint probability
distribution of $\bm{\xi}$. Note that the function, $f$ is considered to be continuous and differentiable
in the considered parameter domain. Hence, it is possible that a given $f(\bm{\xi})$ might not admit an
active subspace. However, it is of remarkable interest to investigate if one exists since the subspace could
be exploited to reduce the dimensionality of the problem and hence the associated computational effort.
$\nabla_{\bm{\xi}}f$ denotes a vector with individual components
being partial derivatives of $f$ with respect to the $i^\text{th}$ input, $\xi_i$. Since $\mat{C}$ is symmetric and
positive semi-definite, it admits a real eigenvalue decomposition:
%
\be
\mat{C} = \bm{W}\bm{\Lambda}\bm{W}^\top
\ee
%
where $\bm{\Lambda}$ = diag($\lambda_1,\ldots,\lambda_{N_p}$) with $\lambda_i$'s
 sorted in descending order, $i:\{1,2,\ldots,N_p\}$. The eigenpairs are partitioned about
 the $i^{\text{th}}$ eigenvalue such that the ratio, $\left(\frac{\lambda_i}{\lambda_{i+1}}\right)$ $\gg$ 1
 as follows:
 %
 \be
 \bm{W} = [\bm{w}_1~\bm{w}_2],~~\bm{\Lambda} = \begin{bmatrix}\bm{\Lambda}_1 & \\  &
  \bm{\Lambda}_2 \\
\end{bmatrix}
 \ee
 %
 where $\bm{w}_1$ comprises the dominant eigenspace of $\mat{C}$ regarded as the active subspace,
 and $\bm{\Lambda}_1$ is the corresponding set of eigenvalues. The active subspace is used to 
 construct the so-called active variables, $\bm{w}_1^\top \bm{\xi}$ and $f(\bm{\xi}(\theta))$ is transformed
 as $G(\bm{w}_1^\top \bm{\xi})$. In practice, the set of samples, $\bm{\xi}_i$'s are projected to the
 physical space as $\bm{\theta}_i$'s at which the model output, $f(\bm{\xi}(\theta))$ is computed. 
 The matrix, $\mat{C}$ is typically approximated using pseudo-random sampling techniques such 
 as Monte Carlo and Latin hypercube sampling. The integral in~\eqref{eq:C} is replaced with a
 summation as follows:
 %
 \be
 \mat{C}\approx \hat{\mat{C}} = \frac{1}{N}\sum\limits_{i=1}^{N} 
 (\nabla_{\bm{\xi}}f(\bm{\xi}_i))(\nabla_{\bm{\xi}}f(\bm{\xi}_i))^\top
 = \hat{\bm{W}}\hat{\bm{\Lambda}}\hat{\bm{W}}^\top
\label{eq:chat}
 \ee
 %
 It has been shown that the accuracy of approximated dominant eigenspace, $\hat{\bm{w}}_1$ is inversely
 proportional to the difference between the smallest eigenvalue in $\hat{\bm{\Lambda}}_1$ and the
 largest eigenvalue in $\hat{\bm{\Lambda}}_2$~\cite{Constantine:2014}.

### 2018-07-27 Paper/method.tex
..

We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$, at an initial set of $n_1$ samples. 
For this purpose, $\bm{\xi}_i$'s are projected to the physical space as $\bm{\theta}_i$'s. Using the gradient
evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition of $\hat{\mat{C}}$ yields an initial
estimate of the dominant eigenspace, $\hat{\bm{W}_1}$ and corresponding set of eigenvalues, $\hat{\bm{\Lambda}_1}$.
At each subsequent iteration, model evaluations are generated at a new set of $n_s$ samples. The new set
of gradient evaluations are augmented with the available set to re-construct $\hat{\mat{C}}$ followed by its eigenvalue
decomposition. The relative
change in the norm of the difference in squared value of individual components of the dominant eigenvectors between subsequent 
iterations is evaluated. The process is terminated and the resulting eigenspace is considered to be converged once
highest relative change at iteration $r$, $\delta w_{1,j}^{(r)}$, is smaller a given tolerance, $\tau$. A regression fit to
 $G(\bm{w}_1^\top\bm{\xi})$ is used as a surrogate to characterize and quantify the uncertainty in the model
 output. Moreover, the components of the dominant eigenvectors are used to compute the activity scores which
 provide an insight into the relative importance of the uncertain parameters. The sequence of steps for the grad-based
 approach are provided in Algorithm~\ref{alg:grad}.

@@

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-based approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $\tau$. 
\Ensure $\Lambda$, $W$, $\eta$. 
    \Procedure{Gradient-based}{}
	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$.
	\State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
	\State $N_\text{total}$ = $n_1$ 
	\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             $k=1, \ldots, N_\text{total}$.  
	\Statex\hspace{5mm} Using Finite Difference:
	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
	\Statex\hspace{5mm} ii. Augment the set of samples with neighboring points.
	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
        \{\xi_{k,j}+d\xi\}_{j=1}^{N_p} \nonumber
	\ee
	\Statex\hspace{5mm} iii. Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$
	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
        to compute $\bm{g}^k$. 
	\State Compute the matrix, $\mathcal{C}$ = 
        $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
        $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$
	\State Set $r$ = 0
	\Loop
		\State Set $r$ = $r$ + 1
		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
		\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
		\State $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
		\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             	$k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
		\State Compute $\mathcal{C}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
		\State Compute $\delta w_{1,j}^{(r)}$ = 
                       \scalebox{1.25}{$\frac{\|(w_{1,j}^{r})^2 - 
                       (w_{1,j}^{r-1})^2\|}{\|(w_{1,j}^{r-1})^2\|}$}, 
                       $j = 1,\ldots,\mathcal{S}$.
		\If {$\max\left(\delta w_{1,j}^{(r)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\eta_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1,j}w_{1,j}^2$,
	$i=1,\ldots,d$.

@@

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-free approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $\alpha$, $M$, $\tau$. 
\Ensure $\Lambda$, $W$, $\eta$. 
    \Procedure{Gradient-free}{}
	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$.
	\State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
	\State $N_\text{total}$ = $n_1$ 
	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
	\State Draw $M$ random samples, $\{\bm{\nu}_k\}_{k=1}^{M}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$.
	\State Construct the matrix, $\mathcal{C}$.
	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
        $\Lambda_1^{(0)}\in \mathbb{R}^{d\times\mathcal{S}}$
	\State Set $r$ = 0
	\Loop
		\State Set $r$ = $r$ + 1
		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
		\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
		\State $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
		\State Compute $G(\bm\theta_k)$, $k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
		\State Construct the matrix, $\mathcal{C}$.
		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
		\State Compute $\delta w_{1,j}^{(r)}$ = 
                       \scalebox{1.25}{$\frac{\|(w_{1,j}^{r})^2 - 
                       (w_{1,j}^{r-1})^2\|}{\|(w_{1,j}^{r-1})^2\|}$}, 
                       $j = 1,\ldots,\mathcal{S}$.
		\If {$\max\left(\delta w_{1,j}^{(r)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\eta_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1,j}w_{1,j}^2$,
	$i=1,\ldots,d$.
    \EndProcedure
  \end{algorithmic}
  \label{alg:free1}
\end{breakablealgorithm}
\bigskip

@@

As mentioned earlier, the model output $f(\bm{\xi}(\bm{\theta}))$ predominantly varies in the active subspace. Hence, 
$f(\bm{\xi}(\bm{\theta}))$ can be transformed as a function of the active variable as $G(\bm{w}_1^\top\bm{\xi})$ in
the subspace.

>>>>

We begin by evaluating the gradient of the model output, $\nabla_{\bm{\xi}}f$, at an initial set of $n_1$ samples. 
For this purpose, $\bm{\xi}_i$'s are projected to the physical space as $\bm{\theta}_i$'s. Using the gradient
evaluations, the matrix, $\hat{\mat{C}}$ is computed. Eigenvalue decomposition of $\hat{\mat{C}}$ yields an initial
estimate of the dominant eigenspace, $\hat{\bm{w}_1}$ and corresponding set of eigenvalues, $\hat{\bm{\Lambda}_1}$.
At each subsequent iteration, model evaluations are generated at a new set of $n_s$ samples. The new set
of gradient evaluations are augmented with the available set to re-construct $\hat{\mat{C}}$ followed by its eigenvalue
decomposition. The relative
change in the norm of the difference in squared value of individual components of the dominant eigenvectors between subsequent 
iterations is evaluated. The process is terminated and the resulting eigenspace is considered to be converged once
highest relative change at iteration $r$, $\delta w_{1,j}^{(r)}$, is smaller a given tolerance, $\tau$. A regression fit to
 $G(\bm{w}_1^\top\bm{\xi})$ is used as a surrogate to characterize and quantify the uncertainty in the model
 output. Moreover, the components of the dominant eigenvectors are used to compute the activity scores which
 provide an insight into the relative importance of the uncertain parameters. The sequence of steps for the grad-based
 approach are provided in Algorithm~\ref{alg:grad}.

@@

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-based approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $\tau$. 
\Ensure $\Lambda$, $W$, $\eta$. 
    \Procedure{Gradient-based}{}
	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$.
	\State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
	\State $N_\text{total}$ = $n_1$ 
	\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             $k=1, \ldots, N_\text{total}$.  
	\Statex\hspace{5mm} Using Finite Difference:
	\Statex\hspace{5mm} i. Assign a small increment, $d\xi$.
	\Statex\hspace{5mm} ii. Augment the set of samples with neighboring points.
	\be \{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}:~\{\bm{\xi}_k\}_{k=1}^{n_1} \cup
        \{\xi_{k,j}+d\xi\}_{j=1}^{N_p} \nonumber
	\ee
	\Statex\hspace{5mm} iii. Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\Xi}_k\}_{k=1}^{n_1(N_p+1)}$
	\Statex\hspace{5mm} iv. Using the augmented set, $\{\bm{\theta}_k\}_{k=1}^{n_1(N_p+1)}$
        to compute $\bm{g}^k$. 
	\State Compute the matrix, $\mathcal{C}$ = 
        $\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
        $\Lambda_1^{(0)}\in \mathbb{R}^{N_p\times\mathcal{S}}$
	\State Set $r$ = 0
	\Loop
		\State Set $r$ = $r$ + 1
		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
		\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
		\State $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
		\State Compute $\bm{g}^k = \nabla_{\bm{\theta}}G(\bm\theta_k)$, 
             	$k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
		\State Compute $\mathcal{C}$ = 
        	$\frac{1}{N_\text{total}}\sum\limits_{k=1}^{N_\text{total}}[\bm{g}^k][\bm{g}^k]^\top$
		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
		\State Compute $\delta w_{1,j}^{(r)}$ = 
                       \scalebox{1.25}{$\frac{\|(w_{1,j}^{r})^2 - 
                       (w_{1,j}^{r-1})^2\|}{\|(w_{1,j}^{r-1})^2\|}$}, 
                       $j = 1,\ldots,\mathcal{S}$.
		\If {$\max\left(\delta w_{1,j}^{(r)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}w_{1_{i,j}}^2$,
	$i=1,\ldots,d$.

@@

\bigskip
\begin{breakablealgorithm}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
  \caption{An iterative gradient-free approach for discovering the active subspace}
  \begin{algorithmic}[1]
\Require $\theta_l$, $\theta_u$, $M$, $\tau$. 
\Ensure $\Lambda$, $W$, $\eta$. 
    \Procedure{Gradient-free}{}
	\State Draw $n_1$ random samples, $\{\bm{\xi}_k\}_{k=1}^{n_1}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$.
	\State Project to the physical space:
        $\{\bm{\theta}_k\}_{k=1}^{n_1}=\theta_l+0.5(\theta_u-\theta_l)\{\bm{\xi}_k\}_{k=1}^{n_1}$
	\State $N_\text{total}$ = $n_1$ 
	\State Compute $G(\bm\theta_k)$, $k=1, \ldots, N_\text{total}$.
	\State Draw $M$ random samples, $\{\bm{\nu}_k\}_{k=1}^{M}$ $\in$ [-1,1]
         according to $\bm{f(\xi)}$.
	\State Construct the matrix, $\mathcal{C}$.
	\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(0)}\Lambda^{(0)} W^{(0)\top}$
	\State Partition the Eigenpairs: $\Lambda^{(0)}~=~ 
        \begin{bmatrix} \Lambda_1^{(0)} & \\ & \Lambda_2^{(0)} \end{bmatrix}$, 
        $W^{(0)}~=~\begin{bmatrix} w_1^{(0)} & w_2^{(0)} \end{bmatrix}$, 
        $\Lambda_1^{(0)}\in \mathbb{R}^{d\times\mathcal{S}}$
	\State Set $r$ = 0
	\Loop
		\State Set $r$ = $r$ + 1
		\State Draw $n_r$ = $\lceil \beta n_1 \rceil$ new random samples 
                $\{\bm{\xi}_k\}$ $\in$ [-1,1], $k = n_{r-1}+1,\ldots,n_{r-1}+n_r$.
		\State Project $\bm{\xi}_k$~$\rightarrow$~$\bm{\theta}_k$.
		\State $N_\text{total}$ = $N_\text{total}$ + $n_r$ 
		\State Compute $G(\bm\theta_k)$, $k=n_{r-1}+1, \ldots, n_{r-1}+n_r$.  
		\State Construct the matrix, $\mathcal{C}$.
		\State Eigenvalue decomposition, $\mathcal{C}$ = $W^{(r)}\Lambda^{(r)} W^{(r)\top}$
		\State Compute $\delta w_{1,j}^{(r)}$ = 
                       \scalebox{1.25}{$\frac{\|(w_{1,j}^{r})^2 - 
                       (w_{1,j}^{r-1})^2\|}{\|(w_{1,j}^{r-1})^2\|}$}, 
                       $j = 1,\ldots,\mathcal{S}$.
		\If {$\max\left(\delta w_{1,j}^{(r)}\right)<\tau$}
			\State break
		\EndIf
	\EndLoop
	\State Compute $\alpha_i(N_\text{total}) = \sum\limits_{j=1}^{\mathcal{S}} \Lambda_{1_{i,j}}w_{1_{i,j}}^2$,
	$i=1,\ldots,d$.
    \EndProcedure
  \end{algorithmic}
  \label{alg:free1}
\end{breakablealgorithm}
\bigskip

@@

As mentioned earlier, the model output $f(\bm{\xi}(\bm{\theta}))$ predominantly varies in the active subspace. Hence, 
$f(\bm{\xi}(\bm{\theta}))$ can be transformed as a function of the active variable as $G(\bm{w}_1^\top\bm{\xi})$ in
the subspace. The plot of $G$ versus $\bm{w}_1^\top\bm{\xi}$, regarded as the \textit{sufficient summary plot} (SSP) is 
compared for the two approaches in Figure~\ref{fig:comp_ssp}.
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.48\textwidth]{./Figures/comp_ssp}
\caption{An illustrative comparison of the SSPs generated using the grad-based and the grad-free approaches.}
\label{fig:comp_ssp}
\end{center}
\end{figure}
%
The two SSPs are observed to be in excellent agreement with each other. Moreover, it is interesting to note
that the variability in ignition
delay based on the considered prior marginals for $A_i$'s is observed to be approximately linear in the
active subspace; consistent with the fact the it is 1-dimensional.

We extend the comparative assessment of the two approaches by estimating the activity score for individual
uncertain parameters ($\alpha_i$) using
the components of the dominant eigenvector in the following equation:
%
\be
\alpha_i(n) = \sum\limits_{j=1}^{\mathcal{S}}\Lambda_{1_{i,j}}w_{1_{i,j}}^2,~i = 1,\ldots,N_p
\ee
%
where $n$ is the sample size. The activity scores for the 19 uncertain pre-exponents ($A_i$'s), estimated using
the two approaches are plotted in Figure~\ref{fig:comp_as} (left). Additionally, for the purpose of verification, we 
compare these estimates with those obtained using the DGSMs~\cite{Vohra:2018} in 
Figure~\ref{fig:comp_as} (right). 
%
\begin{figure}[htbp]
 \begin{center}
  \includegraphics[width=0.45\textwidth]{./Figures/comp_as}
  \includegraphics[width=0.45\textwidth]{./Figures/ub_conv_kinetics_rich}
\caption{Left: A bar-graph of activity scores ($\alpha_i$'s) for the 19 uncertain pre-exponents ($A_i$'s).
Right: A plot of the screening metric ($\widehat{\mathcal{C}_i\mu_i}$), defined as a normalized product of the
Poincare constant and the DGSM ($\mu_i$) (adapted from~\cite{Vohra:2018}).}
\label{fig:comp_as}
\end{center}
\end{figure}
%
The activity scores estimated using the two approaches agree favourably with each other as well as those
based on the screening metric involving the DGSMs from~\cite{Vohra:2018}. It is observed that the uncertainty
associated with the ignition delay is largely due to the uncertainty in $A_9$ and $A_1$. Sensitivity towards
$A_{15}$ and $A_{17}$ is also found to be significant.

The above comparisons establish consistency between the two approaches. Note that the computational
effort in the case of grad-free approach is significantly smaller. Hence, we are encouraged to use the
grad-free approach for the 33-dimensional application in the next section.
